Safe Transferable Regions

### 2015-11-19 paper/introduction.tex
updated intro

This example is an instance of a general pattern, where a producer
creates a data-structure and passes it to a consumer. In a system
where most of the computation takes this form, and these
data-structures are very large, as is the case with many streaming
big-data analysis systems, garbage collection overhead becomes
significant. Importantly, much of this overhead results from the
collector performing avoidable or unproductive
work~\cite{Broom:HotOS}. For example, a consumer process may trigger a
GC as soon as it receives a large data-structure from a producer
process, which has just finished its GC. In such case, consumer's GC
traverses the entire data-structure performing little or no work.
Another example is the process executing the code in
Fig.~\ref{fig:motivating-eg}, whose GC might repeatedly traverse the
\C{map} data-structure, although its objects cannot be collected until
a suitable timing message arrives.
% as the collector scans these large data-structures For instance, in
% the above example, the list at \C{map[t]} cannot be collected until
% the the timing message arrives, and the list is transferred.
% \C{transfer} operation is complete.  using alternative memory
% management techniques.

An important observation, in the context of processes of kind
described above, is that the data-structures exchanged between them
have can be partitioned into sets of fate-sharing objects with common
lifetimes, which makes them good candidates for a region-based memory
management discipline. A region is a block of memory that is allocated
and freed in one shot, consuming constant time. A region may contain
one or more contiguous range of memory locations, and individual
objects may be dynamically allocated within the region over time,
while they are deallocated en masse when the region is freed.  Thus, a
region is a good fit for a set of fate-sharing objects.
% that is used to realize a data-structure with a well-defined
% lifetime. 
In this example, the output to be constructed for each time window $t$
can be implemented using a separate region.

>>>>

This example is an instance of a general pattern, where a producer
creates a data-structure and passes it to a consumer. In a system
where most of the computation takes this form, and these
data-structures are very large, as is the case with many streaming
big-data analysis systems, garbage collection overhead becomes
significant~\cite{Broom:HotOS}.
Furthermore, in a distributed dataflow system, the GC pause at one node can
have a cascading adverse effect on the performance of other nodes.
For instance, a GC pause at an upstream actor can block
downstream actors that are waiting for messages. 
Much of the GC overhead results from the collector performing avoidable or unproductive work.
For example, in the process executing the code from
Fig.~\ref{fig:motivating-eg}, GC might repeatedly traverse the \C{map}
data-structure, although its objects cannot be collected until a
suitable timing message arrives.

% actors.
% performance degradation due to
% GC overhead becomes aggravated in the distributed 
% For example, a consumer process may trigger a
% GC as soon as it receives a large data-structure from a producer
% process, which has just finished its GC. In such case, consumer's GC
% traverses the entire data-structure performing little or no work.

% as the collector scans these large data-structures For instance, in
% the above example, the list at \C{map[t]} cannot be collected until
% the the timing message arrives, and the list is transferred.
% \C{transfer} operation is complete.  using alternative memory
% management techniques.

An important observation, in the context of processes of kind
described above, is that the data-structures exchanged between them
can be partitioned into sets of fate-sharing objects with common
lifetimes, which makes them good candidates for a region-based memory
management discipline. A region is a block of memory that is allocated
and freed in one shot, consuming constant time. A region may contain
one or more contiguous range of memory locations, and individual
objects may be dynamically allocated within the region over time,
while they are deallocated en masse when the region is freed.  Thus, a
region is a good fit for a set of fate-sharing objects.
% that is used to realize a data-structure with a well-defined
% lifetime. 
In this example, the output to be constructed for each time window $t$
can be implemented using a separate region.
